The heuristic algorithm given above for calculating row displacements does not perform well in the worst case.
The intuition behind this drawback is that rows with many nonzero entries block a lot of possible displacements for the remaining rows.
Since the displacements for the rows with many nonzeros are calculated first, the first few rows may block a big chunk and in consequence later rows are assigned larger displacements.

To improve on the previous method, we introduce another layer of indirection:
We will add a set of column displacements $c_i$ that are applied before the row displacements are calculated.
The column displacements transform a table $A$ into a larger table $B$ with more rows and are supposed to distribute the nonnull entries in $B$ better.
Here, better means that the increased size of $B$ makes up for the unfavourable distribution of nonnulls in the original table $A$.

The column displacements $c_i$ are calculated similar to the row displacements.
The first column is assigned a displacement of zero ($c_1 = 0$).
Then, from the left to the right for each column $j$, $c_j$ is increased stepwise until the \emph{exponential decay condition} $E_j$ is satisfied.
Before explaining $E_j$, let $B_j$ be the first $j$ displaced columns of $A$ and $n_j(i)$ be the total number of nonzeros in rows of $B_j$ that have more than $i$ nonzeros.
Then $n_j(i) = 0$ if $i > j$ and $n_j(i)$ is non-increasing in $i$.
$E_j$ then states that for all $i$, $n_j(i)$ must decrease exponentially.
Our improved storage scheme for a table $A$ is now:
First, calculate the column displacements $c_i$ for $A$.
Apply the $c_i$ to $A$.
Since the rows are shifted independently of each other, there may be empty spaces before or after a shifted column.
For example, this can be seen in Figure~\ref{fig:row_disp} where the space left of the third to fourth row and right of the first, second and fourth row is empty.
This empty space is filled with zero entries to create a complete, possibly non square table before the next step.
Note that in Figure~\ref{fig:row_disp}, row displacements have been applied, however it is the same for column displacements.
Next, calculate the row displacements $r_i$ for $B$ and use them to collapse $B$ into an array $C$.

To look up a value $k$, we calculate its position $(i, j)$ as seen above.
After that we apply the row and columns displacements to get an index $k' = r_{i + c_j} + j$ into $C$.
$A$ contained $k$ if $C(k')$ equals $k$.
A few words on the formula for $k'$.
The index $i + c_j$ points to the row where $k$'s cell is moved to by the column displacements.
This row in turn is shifted by its row displacement.
This means that $r_{i + c_j}$ points to the start of the row in which $k$ lies after applying column and row displacements.
To find $k$'s position in $C$, we have to add $j$ to the start of the row and together we get the above formula for $k'$.
